Active Learning : From Blackbox To Timed Connectors

### 2016-01-09 paper/paper.tex
.

\begin{document}
\maketitle
\begin{abstract}
  Coordination is becoming more and more important, especially in concurrent programming and
  distributed systems. More and more tools are built to verify coordination models while most of
  them are facing the same problem: \emph{lack of models}. In many cases, somehow we need to check a
  connector with nothing more than a binary file. In this paper, we present a learning-based
  framework to formally verify a Reo connector in case there is no source code. From some practical
  cases, we've shown the effciency of our approach. Both the algorithm and Reo itself have been
  ported to \emph{Golang}, making it fully-prepared for parallel programming.
\end{abstract}

\begin{IEEEkeywords}
  active learning, coordination, formal method.
\end{IEEEkeywords}

@@

Now we can use Parameterized Mealy Machines to define a new semantics to Reo.
\begin{example}[PMM Semantics of FIFO channel $PM_{FIFO}$]
  The semantics of FIFO channel with source end $A$ and sink end $B$ can be defined as follows.
  \begin{itemize}
    \item[-] $S(\Sigma)=\{q_0\}\cup\{q_d|d\in\Sigma\}$
    \item[-] $I=\{A\}$
    \item[-] $O=\{B\}$
    \item[-] $s_0=q_0$
    \item[-] output function
      \begin{displaymath}
        \delta(\Sigma)(s,i)=\left\{
        \begin{array}[h]{ll}
          (B:\bot) & s=q_0\land i=(A:\_,B:\rnoread) \\
          (B:d) & s=q_0\land i=(A:d,B:\rread)\\
          \rblock & s=q_0\land i=(A:\bot,B:\rread) \\
          (B:d) & s=q_d\land i=(A:\_,B:\rread)\\
          \rblock & s=q_d\land i=(A:d,B:\rnoread) \\
          (B:\bot) & s=q_d\land i=(A:\_,B:\rnoread) \\     
        \end{array}
        \right.
      \end{displaymath}
    \item[-] transition function
      \begin{displaymath}
        \lambda(\Sigma)(s,i)=\left\{
        \begin{array}[h]{ll}
          q_d & s=q_0\land i=(A:d,B:\rnoread) \\
          q_d & s=q_{d'}\land i=(A:d,B:\rread) \\
          q_0 & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
  \end{itemize}
\end{example}

@@

As shown in Table \ref{tabel:timecost}, time comsumption mainly comes from membership queries.
Since time is involved in our model, it's inevitably that simulation takes time to behave normally.
Even worse, since our models are treated as blackbox, there's little room for optimization. The only
way to reduce time cost is trying not to call the simulation process. Basic ideas are listed as
follows:

@@

To reduce these repeated simulation, we use a cache to save all simulation results. 
\begin{table}[h]
  \renewcommand{\arraystretch}{1.3}
  \caption{Reduction of Membership Queries}
  \label{tabel:cacheoptimization}
  \centering
  \begin{tabular}{l||rrr}
    \hline
    & FIFO & Alternator & Gate \\
    \hline\hline
    Before & 41.571 &  & 1034 \\
    After & 41.571 &  & 707 \\
    Reduction Rate & & & 31.6\% \\
    \hline
  \end{tabular}
\end{table}

>>>>

\begin{document}
\maketitle
\begin{abstract}
  % reo coordination models
  Coordination languages provide a mean to formally specify the communication and
  reaction between different components. Such models play key roles in concurrent programming and
  large-scale distributed systems.
  % algorithms
  In this paper, an active learning framework is proposed to extract timed connectors from black-box
  systems.
  First of all, we introduce parameterized mealy machine as an operational semantics for
  channel-based coordination language Reo.
  Parameterized mealy machine serves as a bridge between Reo connectors and mealy machines. With
  production operator, complex connectors can be composed by basic channels and transformed into
  mealy machines. 
  Moreover, we adapt L*, a well-known learning algorithm, to timed connectors (in form of mealy
  machines). The new algorithm has shown its efficiency in multiple cases.
  % implementation
  Implementations of this framework is provided as a package in \texttt{Golang}.
\end{abstract}

\begin{IEEEkeywords}
  Active Learning, Coordination Languages, Formal Methods
\end{IEEEkeywords}

@@

Now we can use Parameterized Mealy Machines to define a new semantics to Reo.
\begin{example}[PMM Semantics of FIFO channel $PM_{FIFO}$]
  \label{example:pmmfifo}
  The semantics of FIFO channel with source end $A$ and sink end $B$ can be defined as follows.
  \begin{itemize}
    \item[-] $S(\Sigma)=\{q_0\}\cup\{q_d|d\in\Sigma\}$
    \item[-] $I=\{A\}$
    \item[-] $O=\{B\}$
    \item[-] $s_0=q_0$
    \item[-] output function
      \begin{displaymath}
        \delta(\Sigma)(s,i)=\left\{
        \begin{array}[h]{ll}
          (B:\bot) & s=q_0\land i=(A:\_,B:\rnoread) \\
          (B:d) & s=q_0\land i=(A:d,B:\rread)\\
          \rblock & s=q_0\land i=(A:\bot,B:\rread) \\
          (B:d) & s=q_d\land i=(A:\_,B:\rread)\\
          \rblock & s=q_d\land i=(A:d,B:\rnoread) \\
          (B:\bot) & s=q_d\land i=(A:\_,B:\rnoread) \\     
        \end{array}
        \right.
      \end{displaymath}
    \item[-] transition function
      \begin{displaymath}
        \lambda(\Sigma)(s,i)=\left\{
        \begin{array}[h]{ll}
          q_d & s=q_0\land i=(A:d,B:\rnoread) \\
          q_d & s=q_{d'}\land i=(A:d,B:\rread) \\
          q_0 & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
  \end{itemize}
\end{example}

@@

As shown in Table \ref{tabel:timecost}, time consumption mainly comes from membership queries.
Since time is involved in our model, it's inevitably that simulation takes time to behave normally.
Even worse, since our models are treated as blackbox. With no access to inner behaviour of the
connectors, it's almost impossible to accelerate the simulation process.

Fortunately, there are still other optimization solutions. After reviewing our algorithm, we found
that simulations on similar sequences were invoked frequently:

@@

If previous simulation results are stored in a well-maintained cache, the time-cost in
simulation process could be reduced signficiantly. In this work, we use a multiway tree to buffer
these results. A brief example of such trees can be found in Example \ref{example:tree}.

\begin{example}[Multiway-Tree Cache]
  \label{example:tree}
  Considering the FIFO channel presented in Example \ref{example:pmmfifo}. 
  \todo{draw a graph for multiway-tree cache}
\end{example}

\begin{table}[h]
  \renewcommand{\arraystretch}{1.3}
  \caption{Reduction of Membership Queries}
  \label{tabel:cacheoptimization}
  \centering
  \begin{tabular}{l||rrr}
    \hline
    & FIFO & Alternator & Gate \\
    \hline\hline
    Before & 93 & 880 & 1034 \\
    After & 90 & 725 & 707 \\
    Reduction Rate & 3.2\% & 21.4\% & 31.6\% \\
    \hline
  \end{tabular}
\end{table}
